---
title: "Labor Rights and Global Supply Chains"
subtitle: "An Annotated Bibliography"
geometry: margin = 1in
output: 
  pdf_document:
    citation_package: biblatex
    # citation_package: natbib
biblatexoptions: [backend = biber, skipbib = true, labeldateparts = true, isbn = false, annotation = false, url = false]
#biblatexoptions: [backend = biber, skipbib = true, pagetracker=true, autocite=inline, alldates=comp, labeldateparts=true, citetracker=true, uniquename=minfull, useeditor=true, usetranslator=true, usenamec=true, alltimes=12h, urltime=24h, datecirca=true, datezeros=false, dateuncertain=true, timezones=true, compressyears=true, ibidtracker=constrict, sorting=cms, punctfont, cmslos=true, nodates, uniquelist=minyear, maxbibnames=10, minbibnames=7, sortcase=false, abbreviate=false, dateabbrev=false, avdate=true]
# biblio-style: chicago-notes
biblio-style: chicago-authordate
# citation-style: 
bibliography: labor-rights.bib
urlcolor: blue
fontsize: 12pt
header-includes: 
  - \usepackage[american]{babel}
  # - \usepackage{doi}
  # - \usepackage[backend=biber, style = chicago-authordate]{biblatex}
  # - \usepackage{texlive-bibtex-extra}
  # - \usepackage{chicago}
  # - \usepackage{natbib}
  # - \usepackage{bibentry}
  # - \bibliographystyle{plain}
  # - \nobibliography{labor-rights.bib}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, results="asis")
library(RefManageR)
bib <- ReadBib("labor-rights.bib", check=FALSE)
#library(rcrossref)
#library(bibtex)
#library(bib2df)
#bib <- bib2df("labor-rights.bib")
##apa_style <- tools::bibstyle("apa")
#myopts <- BibOptions(bib.style = "authoryear", style = "latex", first.inits=FALSE, max.names = 20)
```

### Can you change the citation style and make the font a little bigger? 
### Garbled text in some abstracts. How to fix? 
### Create a template for annotated bib? (Study syllabus template, alter to include Chicago style, post) 
### This is pretty nice. Only thing is: what happens when I have a web-based article that I want to include, since I disabled urls? Will it still show the urls for websites? 
### Also, do I want to keep the ISBN for books (but delete ISSN for articles?) Would that provide a link to the book somehow? 
### Not able to get a space between references anymore using \\; investigate \break, \newline and \bigbreak.

\fullcite{Anner2017} 

```{r Anner_abstract}
cat("Abstract: ")
cat(bib["Anner2017"]$abstract)
```

Notes: Now I would like to write some notes on this article.

```{r crossref_test}
#cr_cn(dois = "10.1111/1758-5899.12385", format = "text", style = "american-political-science-association")
#anner_2017 <- cr_cn(dois = "10.1111/1758-5899.12385", format = "text", style = "american-political-science-association")
#cat(anner_2017)
```

